Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)

### 2018-03-12 paper.tex
update

\subsubsection{Allocating Addresses}
Up to now, we did not describe much about allocating adresses in stores and the 
time-stamp @Time@.
But in abstract interpretation, they are key ingredients to achieve 
analyses with different sensitivities, as well as to a finite state space analysis\cite{Gilray:2016:ACP:2951913.2951936}.
To effectively approximate the runtime behavior, a finite 
program contour @time@ that encodes the program execution history is introduced.
Here, we use a list of execution contexts (expressions) to represent that, and
as we will see in section \ref{kcfainst}, by using different @tick@
function on time we may obtain a family of analyses.

@@

As we mentioned previously, the space of addresses has to be finite in AAM.
Binding addresses are parameterized by variable names and the time that creates this binding,
which are both finite.
Continuation addresses @KAddr@ has two variants: 
1) @Halt@ which corresponds to the empty stack, and
2) @ContAddr@ consists of the entry expressions of callee, which are also also
finite.
One may ask, why not also keep track of the time in continuation addresses?
(TODO: why target expression, allocation polyvariance)

@@

\subsubsection{$k$-call-sensitive Instantiation} \label{kcfainst}

@@

The state transition function @step@ is still non-deterministic, but
the only non-determinism happens when dereferencing the callee @f@ 
from function applications @App(f, ae)@.

@@

@f1@ and @f2@ mutually call each others, and the stack will be alternatively
pushing @f1@ and @f2@ onto the top of current stack, but not existing 
two stack components are identical in the state space.

@@

type MCont = VS => Set[VS]
type Cont  = MCont => (VS => Set[VS])

def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont, rcont: MCont): Set[VS] = {
  val new_time = (e::time).take(k)
  e match {
    case Let(x, App(f, ae), e) =>
      val closures = atomicEval(f, env, store).toList.asInstanceOf[List[Clos]]

      def nd(closures: List[Clos]): Set[VS] = {
        val Clos(Lam(v, body), c_env)::rest = closures
        val baddr = allocBind(v, new_time)
        val new_env = c_env + (v -> baddr)
        val new_store = store.update(baddr, atomicEval(ae, env, store))
        val new_cont = (r: MCont) => (vs: VS) => {
          val baddr = allocBind(x, vs.time)
          val new_env = env + (x -> baddr)
          val new_store = vs.store.update(baddr, vs.vals)
          aval(e, new_env, new_store, vs.time, cont, r)
        }
        val new_rcont = if (rest.isEmpty) {
          rcont 
        } else (vs: VS) => {
          nd(rest) + vs
        }
        aval(body, new_env, new_store, new_time, new_cont, new_rcont)
      }

      nd(closures)

    case ae if isAtomic(ae) =>
      val ds = atomicEval(ae, env, store)
      cont(rcont)(VS(ds, new_time, store))
  }
}

def analyze(e: Expr) = aval(e, Map(),
                            Store[BAddr, Storable](Map()), List(),
                            (r: MCont) => (vs: VS) =>  r(vs),
                            (vs: VS) => Set(vs))
\end{lstlisting}

>>>>

\subsubsection{Allocating Addresses}
Up to now, we did not describe much about allocating adresses in stores and the 
time-stamp @Time@.
But in abstract interpretation, they are key ingredients to achieve 
analyses with different sensitivities, as well as to a finite state space analysis\cite{Gilray:2016:ACP:2951913.2951936}.
To effectively approximate the runtime behavior, a finite 
program contour @time@ that encodes the program execution history is introduced.
Here, we use a list of execution contexts (expressions) to represent that, and
as we will see in section \ref{kcfainst}, by applying different @tick@
functions to time we may obtain a family of analyses.

@@

As we mentioned previously, the space of addresses has to be finite in AAM.
Binding addresses are parameterized by variable names and the time when creating this binding,
which are both finite.
Continuation addresses @KAddr@ has two variants: 
1) @Halt@ which corresponds to the empty stack, and
2) @ContAddr@ consists of the entry expressions of callee, which are also also
finite.
One may ask, why not also keep track of the time in continuation addresses?
(TODO: why target expression, allocation polyvariance)

@@

\subsubsection{$k$-Call-Sensitive Instantiation} \label{kcfainst}

@@

The state transition function @step@ shows in below which is still 
non-deterministic, but the only non-determinism happens when dereferencing the callee @f@ 
from function applications @App(f, ae)@.

@@

Function @f1@ and @f2@ mutually invokes each others, and the stack will be 
alternatively pushing frame @f1@ and @f2@ onto the top of current stack, 
but not existing two stack components are identical in the state space.

@@

type Ans = Set[VS]
type NCont = VS => Ans
type Cont = VS => (NCont => Ans)

def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont, ncont: NCont): Ans = {
  val new_time = (e::time).take(k)
  e match {
    case Let(x, App(f, ae), e) =>
      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]

      def nd(closures: List[Clos]): Ans = {
        val Clos(Lam(v, body), c_env)::rest = closures
        val baddr = allocBind(v, new_time)
        val new_env = c_env + (v -> baddr)
        val new_store = store.update(baddr, aeval(ae, env, store))
        val new_cont = (vs: VS) => (r: NCont) => {
          val baddr = allocBind(x, vs.time)
          val new_env = env + (x -> baddr)
          val new_store = vs.store.update(baddr, vs.vals)
          aval(e, new_env, new_store, vs.time, cont, r)
        }
        val new_ncont = if (rest.isEmpty) {
          ncont
        } else (vs: VS) => {
          nd(rest) + vs
        }
        // new_time is determined by the lexical scope
        aval(body, new_env, new_store, new_time, new_cont, new_ncont)
      }

      nd(closures)
    case ae if isAtomic(ae) =>
      val ds = aeval(ae, env, store)
      cont(VS(ds, new_time, store))(ncont)
  }
}

def mtEnv = Map[String, BAddr]()
def mtStore = Store[BAddr, Storable](Map())
def analyze(e: Expr) = aval(e, env, store, List(), 
                            (vs: VS) => (r: NCont) => r(vs), 
                            (vs: VS) => Set(vs))

\end{lstlisting}
